---
title: "class_3"
author: "Milan"
date: "2023-12-13"
output: html_document
---

Load our libraries
```{r}
library(haven)
library(tidyverse)
```

Read in datasets
```{r}
ads_data <- read_sav("./ads_raw.sav")
ces_2019_raw <- read_csv("./ces_2019_raw.csv")
```


```{r}
glimpse(ads_data)
```
```{r}
filter(ads_data, Duration__in_seconds_ < 100 )
```
```{r}
arrange(ads_data, Duration__in_seconds_)
```
Select
```{r}
select(ads_data, RecordedDate)
```

Multiple Operations
```{r}
ads_data %>%
  filter(Duration__in_seconds_ < 100) %>%
  arrange(Duration__in_seconds_) %>%
  select(RecordedDate,Duration__in_seconds_)
```

```{r}
ads_data <- ads_data %>%
  mutate(Birthyear_add_day =
  str_c(Birthyear, "07-01")) %>%
  mutate(Birthyear_add_day =
  as_datetime(Birthyear_add_day))
```

```{r}
ads_data %>%
  mutate(age = EndDate - Birthyear_add_day)
```
```{r}
summary(ads_data)
```
```{r}
#MEDIAN
ads_data %>%
  pull(Duration__in_seconds_) %>%
  median(na.rm = TRUE)
#MEAN
ads_data %>%
  pull(Duration__in_seconds_) %>%
  mean(na.rm = TRUE)
#RANGE
ads_data %>%
  pull(Duration__in_seconds_) %>%
  range(na.rm = TRUE)
#VARIANCE
ads_data %>%
  pull(Duration__in_seconds_) %>%
  var(na.rm = TRUE)
#STANDARDDEVATION 
ads_data %>%
  pull(Duration__in_seconds_) %>%
  sd(na.rm = TRUE)
```
```{r}
ads_data %>%
  summarise(mean_time =
  mean(Duration__in_seconds_, na.rm = TRUE),
  sd_time = 
    sd(Duration__in_seconds_,na.rm = TRUE)
          )
```

```{r}
ads_data %>%
group_by(Gender) %>%
summarise(count = n(),
mean_time = mean(Duration__in_seconds_, na.rm = TRUE),
sd_time = sd(Duration__in_seconds_, na.rm = TRUE))
```

Data cleaning
```{r}
CES_data <- ces_2019_raw %>%
  mutate(cps19_yob_fix = cps19_yob + 1919)

CES_data %>%
  pull(cps19_yob_fix) %>%
  range(na.rm = TRUE)
```

```{r}
CES_data <- CES_data %>%
  mutate(age = 2019 - cps19_yob_fix)

CES_data %>%
  pull(age) %>%
  range(na.rm = TRUE)

CES_data <- CES_data %>%
  mutate(cps19_gender_fix = 
  factor(cps19_gender)) %>%
  mutate(cps19_gender_fix = 
           fct_recode(
             cps19_gender_fix,
             "M" = "1",
             "F" = "2",
             "NB" = "3"
           ))

CES_data %>%
  select(cps19_gender,cps19_gender_fix)
```

```{r}

CES_data %>%
  filter(cps19_household > 10) %>%
  arrange(-cps19_household) %>%
  pull(cps19_household)
```
```{r}
CES_data <- CES_data %>%
  mutate(cps19_household =
ifelse(cps19_household > 15, NA, cps19_household ))

CES_data %>%
  filter(cps19_household > 10) %>%
  pull(cps19_household)
```

```{r}
CES_data %>%
  filter(cps19_income_number > 10000000) %>%
  arrrange(-cps19_income_number) %>%
  pull(cps19_income_number)
```

```{r}

CES_data <- CES_data %>%
  mutate(cps19_income_number = 
ifelse(cps19_income_number >= 10000000,
       NA, cps19_income_number))

CES_data %>%
  filter(cps19_income_number > 1000000) %>%
  pull(cps19_income_number)
```

Data Summarizing
```{r}

CES_data <- ces_data_cleaned
CES_data %>%
  filter(cps19_province == 22) %>%
  select(cps19_prov_gov_sat,
         cps19_prov_id,
         cps19_income_number)
```

CES_data <- read_csv("./ces_2019_cleaned.csv")

EXCERISE
```{r}
#FILTER WHERE AGE IS BETWEEN 30 & 50
CES_data_filtered <- CES_data %>%
  filter(cps19_age >= 30 & cps19_age <= 50)

CES_data_filtered
```


```{r}
#FILTER data where 
CES_data_filtered <- CES_data %>%
  filter(!is.na(cps19_votechoice))

CES_data_filtered
```

```{r}
CSE_data_selected <- CES_data %>% 
  select(cps19_age,cps19_province)

CSE_data_selected
```

```{r}
CSE_data_selected <- CES_data %>%
  select(-cps19_province)

CSE_data_selected
```
5
```{r}
CES_data <- CES_data %>%
  mutate(news_consumption = 
           ifelse(cps19_news_cons == "0 minutes", "No", "Yes"))

CES_data %>%
  select(news_consumption)
```

6
```{r}
CES_data <- CES_data %>%
mutate(cps19_income_number = cps19_income_number / 1000)

CES_data %>%
  filter(!is.na(cps19_income_number)) %>%
  select(cps19_income_number)
```
CES_data <- read_csv("./ces_2019_cleaned.csv")

```{r}
CES_data_rating <- CES_data %>%
  filter(!is.na(cps19_lead_rating_23)) %>% 
  group_by(cps19_votechoice) %>% 
  summarise(median_rating = median(cps19_lead_rating_23 , na.rm = TRUE),
            mean_rating = mean(cps19_lead_rating_23, na.rm = TRUE))

CES_data_rating  
```
7

```{r}
CES_data_count <- CES_data %>% 
  group_by(cps19_imm,cps19_spend_educ) %>% 
  summarise(count = n())

CES_data_count
```





